Towards a Formally Verified EVM in Production Environment

### 2020-04-25 COORDINATION2020/paper5.tex
paper6 update

\author{Xiyue Zhang\inst{1} \and
Yi Li\inst{1} \and
Meng Sun\inst{1}
}
\authorrunning{X.~Zhang, Y.~Li and M.~Sun}

@@

\begin{abstract}
Among dozens of decentralized computing platforms, Ethe\-reum attracts widespread attention for its native support of smart contracts by means of a virtual machine called EVM. Programs can be developed in various front-end languages. For example, Solidity can be deployed to the blockchain in the form of compiled EVM opcodes. However, such flexibility leads to critical safety challenges. In this paper, we formally define the behavior of EVM in Why3, a platform for deductive program verification, which facilitates the verification of different properties. Furthermore, the extracted implementation in OCaml can be directly integrated into the production environment and tested against the standard test suite. The combination of proofs and testing in our framework serves as a powerful analysis basis for EVM and smart contracts.

@@

However, EVM and smart contracts are faced with several security vulnerabilities. In fact, a lot of attacks on smart contracts succeeded in the past years, such as the famous DAO (Decentralized Autonomous Organization) attack in 2016 resulting in the lost of more than 60M USD, and the Parity Wallet Hacks in which vulnerabilities in the wallet was exploited by an attacker to successfully steal over 150,000 ETH (about 30M USD), and leads to more than 250M USD frozen later in 2017. A taxonomy of vulnerabilities and related attacks against Solidity, the EVM, and the blockchain is presented in \cite{atzei2016survey}.

%The adoption of formal methods can facilitate the production of trustworthy and reliable software systems. Verification such as theorem proving techniques and testing have become essential to guarantee the correctness of safety-critical systems.

%Developing trustworthy smart contracts is a very challenging task due to the complexity of the distributed execution environment and programming languages. In fact, a lot of attacks on smart contracts succeeded in the past years, such as the famous DAO (Decentralized Autonomous Organization) attack in 2016 resulting in the lost of more than 60M USD, and the Parity Wallet Hacks in which vulnerabilities in the wallet was exploited by an attacker to successfully steal over 150,000 ETH (about 30M USD), and leads to more than 250M USD frozen later in 2017.

Formal verification techniques such as theorem proving and testing have become essential to guarantee the correctness of systems. The adoption of formal methods can facilitate the production of trustworthy and reliable smart contracts as well. 
To deal with the security challenges against EVM, we present a formal framework of generating verified EVM for production environment in this paper. The contributions of this work are: 
\begin{itemize}
\item A formal definition of EVM specified in WhyML, the programming and specification language used in Why3 \cite{filliatre2013why3}. 
\item An implementation of EVM in OCaml generated through an extraction mechanism based on a series of customized drivers. 
\item The verification of sample properties and testing of the OCaml implementation for EVM against a standard test suite for Ethereum.
\end{itemize}
%In this paper we propose a framework for modeling and verification of smart contracts based on the Mediator language \cite{LS17} and the UPPAAL model checker \cite{BDLPY11}. There are some existing works for formal verification of smart contracts. \cite{AR18} provides a SMT based approach for verification of Solidity smart contracts. ...
%Comparing with the existing works, the new approach: benefits of Mediator for modeling smart contracts, automatic verification by translation to UPPAAL, probabilistic and statistical model checking ...

This paper is organized as follows: %Section \ref{Sec: Pre} presents some background about Why3, Ethereum Virtual Machine (EVM) and smart contracts. 
We outline the framework for formalizing, property verifying and testing of EVM in Section \ref{Sec: Framework}. %Based on the framework, we perform some experiments in Section \ref{Sec: Experiment} and highlight the verification and testing results for the properties of EVM. 
Section \ref{Sec: Related} presents some related work. Finally, we summarized this paper and point out the future research direction in Section \ref{Sec: Conclusion}.

@@

To formalize the infrastructure of EVM, we first need to provide the formalization of some commonly-used types in EVM, such as the types of machine word and the address in EVM. Hence, we developed a series of type modules such as \texttt{UInt256} and \texttt{UInt160} to ease the representation of corresponding types in EVM. We also use type alias supported by Why3 to make the basic formalization more readable and consistent with the original definition.

@@

The organization of the instruction category is a bit different from the yellow paper~\cite{wood2014ethereum}. The comparison operations are grouped in the arithmetic operation set \texttt{Arith}; and the comparison operations for signed arithmetic are included in the signed arithmetic operation set \texttt{Sarith}. The bitwise operations and operation \texttt{BYTE} are categorized into the bit-related instruction group \texttt{Bits}. The information related instructions including environmental and block information ones are defined in type \texttt{info\_inst}, except \texttt{CALL} and \texttt{CODE} instructions, such as \texttt{CALLDATACOPY}, \texttt{CODECOPY}, \texttt{CALLDATALOAD} and etc. These instructions are more closely related to memory and stack status. Therefore, extra instructions but closely-related ones are added to the original memory and stack instruction groups presented in the yellow paper. In case when some illegal command occurs, the instruction \texttt{Invalid} is included in the \texttt{instruction} definition. The \texttt{STOP} operation and the system operations from the yellow paper are defined in the \texttt{System} instruction set. The specification of the remaining instruction groups are basically the same as the corresponding instruction subsets in \cite{wood2014ethereum}.

@@

Pop and push operations are most commonly-used manipulations for the state transition of stack. Auxiliary functions \texttt{push\_stack} and \texttt{pop\_stack} are defined to control the change of stack status. For capturing the status transition result of \texttt{Swap} instructions, recursive functions \texttt{fetch} and \texttt{drop} are defined to support more direct manipulation on lists. Function \texttt{swap\_stack} is defined further for state transition of swap instructions.
With the pre-defined auxiliary functions, the definition of the interpreter function is essentially comprised of machine status update with regard to concrete instructions.

\subsection{Running EVM in Production Environment}
%We customizedly defined a series of drivers to %extract WhyML programs into Ocaml programs. 
In this section, we introduce the second phase of the framework: deploy the verified EVM, encoded in Why3, in production environments. The deployment is essentially based on a co-compilation framework between OCaml and Rust.

@@

The framework provides the interaction mechanism between Rust, the productive programming language, and Why3 the formal specification language. By gluing them together, verified models can be directly executed in production environments for further testing.

\section{Related Work}\label{Sec: Related}
Research interest of blockchain technology has exploded since the inception  of Bitcoin. As the popularity of the second generation of blockchain, Ethereum, grows, a series of security vulnerabilities have also appeared. Since EVM and smart contracts deal directly with the transactions of valuable c
ryptocurrency units among multiple parties, the security of smart contract programs and EVM implementations is of paramount importance. To address the security challenges, researchers resorted to the techniques of formal methods and program analysis.

\begin{itemize}
    \item \textbf{Formalization Foundation of EVM} An executable formal semantics of EVM has been created in K framework by Everett et al.~\cite{hildenbrandt2017kevm}. Hirai~\cite{hirai2017defining} proposed an EVM implementation in Lem, a language that can be compiled for a few interactive theorem provers. Further, safety properties of smart contracts can be proved in proof assistants like Isabelle/HOL. The \texttt{hevm} project~\cite{hevm} is implemented in Haskell for unit testing and debugging smart contracts, though the EVM implementation is not completed yet. Besides the above formalizations of EVM, there are also implementations of EVM in Javascript~\cite{vmjs}, Go~\cite{vmgo}, Ruby~\cite{cryptape} and etc.
    \item \textbf{Verification of Smart Contracts} Sergey et al.~\cite{sergey2017concurrent} provided a new perspective between smart contracts and concurrent objects, based on which existing tools and insights for understanding, debugging and verifying concurrent objects can be used on smart contract behaviors. In \cite{luu2016making}, several new security problems were pointed out and a way to enhance the operational semantics of Ethereum was proposed to make smart contracts less vulnerable. Due to the difficulty of correcting the semantics of Ethereum, Luu et al.~\cite{luu2016making} also implemented a symbolic execution tool \texttt{OYENTE} to find security bugs. A framework to analyze and verify both the runtime safety and the functional correctness of Solidity contracts in F* was presented in \cite{bhargavan2016formal}.
\end{itemize}

\section{Conclusion and Future Work}\label{Sec: Conclusion}

>>>>

\author{Xiyue Zhang,
Yi Li and
Meng Sun$^*$
}
\authorrunning{X.~Zhang, Y.~Li and M.~Sun}

@@

\begin{abstract}
Among dozens of decentralized computing platforms, Ethe\-reum attracts widespread attention for its native support of smart contracts by means of a virtual machine called EVM. Programs can be developed in various front-end languages. For example, Solidity can be deployed to the blockchain in the form of compiled EVM opcodes. However, such flexibility leads to critical safety challenges. In this paper, we formally define the behavior of EVM in Why3, a platform for deductive program verification, which facilitates the verification of different properties. The extracted implementation in OCaml can be directly integrated into the production environment and tested against the standard test suite. The combination of proofs and testing in our framework serves as a powerful analysis basis for EVM and smart contracts.

@@

However, EVM and smart contracts are faced with several security vulnerabilities. %In fact, a lot of attacks on smart contracts succeeded in the past years, such as the famous DAO (Decentralized Autonomous Organization) attack in 2016 resulting in the lost of more than 60M USD, and the Parity Wallet Hacks in which vulnerabilities in the wallet was exploited by an attacker to successfully steal over 150,000 ETH (about 30M USD), and leads to more than 250M USD frozen later in 2017. 
A taxonomy of vulnerabilities and related attacks against Solidity, the EVM, and the blockchain is presented in \cite{atzei2016survey}. 
%Formal verification techniques such as theorem proving and testing have become essential to guarantee the correctness of systems. The adoption of formal methods can facilitate the production of trustworthy and reliable smart contracts as well. 
To deal with the security challenges against EVM, we developed a formal framework of generating verified EVM for production environment in this paper. The contributions of this work are: 
\begin{itemize}
\item A formal definition of EVM specified in WhyML, the programming and specification language used in Why3 \cite{filliatre2013why3}. 
\item An implementation of EVM in OCaml generated through an extraction mechanism based on a series of customized drivers. 
\item The verification of sample properties and testing of the OCaml implementation for EVM against a standard test suite for Ethereum.
\end{itemize}
%In this paper we propose a framework for modeling and verification of smart contracts based on the Mediator language \cite{LS17} and the UPPAAL model checker \cite{BDLPY11}. There are some existing works for formal verification of smart contracts. \cite{AR18} provides a SMT based approach for verification of Solidity smart contracts. ...
%Comparing with the existing works, the new approach: benefits of Mediator for modeling smart contracts, automatic verification by translation to UPPAAL, probabilistic and statistical model checking ...

This paper is organized as follows: %Section \ref{Sec: Pre} presents some background about Why3, Ethereum Virtual Machine (EVM) and smart contracts. 
We outline the framework for formalizing, property verifying and testing of EVM in Section \ref{Sec: Framework}. %Based on the framework, we perform some experiments in Section \ref{Sec: Experiment} and highlight the verification and testing results for the properties of EVM. 
Section \ref{Sec: Related} presents some related work. Finally, we summarized this paper %and point out the future research direction 
in Section \ref{Sec: Conclusion}.

@@

To formalize the infrastructure of EVM, we need to provide the formalization of commonly-used types in EVM, such as the types of machine word and the address in EVM. Hence, we developed a series of type modules such as \texttt{UInt256} and \texttt{UInt160} to ease the representation of corresponding types in EVM. Type alias supported by Why3 are also used to make the basic formalization more readable and consistent with the original definition.

@@

The organization of the instruction category is a bit different from the yellow paper~\cite{wood2014ethereum}. The comparison operations are grouped in the arithmetic operation set \texttt{Arith}; and the comparison operations for signed arithmetic are included in the signed arithmetic operation set \texttt{Sarith}. The bitwise operations and operation \texttt{BYTE} are categorized into the bit-related instruction group \texttt{Bits}. The information related instructions including environmental and block information ones are defined in type \texttt{info\_inst}, except \texttt{CALL} and \texttt{CODE} instructions, such as \texttt{CALLDATACOPY}, \texttt{CODECOPY} and \texttt{CALLDATALOAD}. These instructions are more closely related to memory and stack status. Therefore, extra instructions but closely-related ones are added to the original memory and stack instruction groups. % presented in the yellow paper. 
In case when some illegal command occurs, the instruction \texttt{Invalid} is included in the \texttt{instruction} definition. The \texttt{STOP} operation and the system operations from the yellow paper are defined in the \texttt{System} instruction set. The specification of the remaining instruction groups are basically the same as the corresponding instruction subsets in \cite{wood2014ethereum}.

@@

Pop and push operations are most commonly-used manipulations for the state transition of stack. Auxiliary functions \texttt{push\_stack} and \texttt{pop\_stack} are defined to control the change of stack status. For capturing the status transition result of \texttt{Swap} instructions, recursive functions \texttt{fetch} and \texttt{drop} are defined to support more direct manipulation on lists. Function \texttt{swap\_stack} is defined further for state transition of swap instructions.
With these pre-defined auxiliary functions, the definition of the interpreter function is essentially comprised of machine status update with regard to concrete instructions.

\subsection{Running EVM in Production Environment}
%We customizedly defined a series of drivers to %extract WhyML programs into Ocaml programs. 
We now introduce the second phase of the framework: deploy the verified EVM, encoded in Why3, in production environments. The deployment is essentially based on a co-compilation framework between OCaml and Rust.

@@

The framework provides the interaction mechanism between Rust and Why3. By gluing them together, verified models can be directly executed in production environments for further testing. The coupling between Rust and extracted OCaml implementation enables us to perform VM tests to test the basic workings of the verified VM. Information of the overarching environment is obtained through the interface of Rust implementation. The test can be performed on the execution of the OCaml implementations to check the operations in different transactions.

\section{Related Work}\label{Sec: Related}
Research interest of blockchain technology has exploded since the inception of Bitcoin. As the popularity of the second generation of blockchain, Ethereum, grows, a series of security vulnerabilities have also appeared. Since EVM and smart contracts deal directly with the transactions of valuable cryptocurrency units among multiple parties, the security of smart contracts and EVM implementations is of paramount importance. To address the security challenges, researchers resorted to the techniques of formal methods and program analysis.

%\begin{itemize}
%    \item \textbf{Formalization Foundation of EVM} 
An executable formal semantics of EVM has been created in K framework by Everett et al.~\cite{hildenbrandt2017kevm}. Hirai~\cite{hirai2017defining} proposed an EVM implementation in Lem, a language that can be compiled for a few interactive theorem provers. Further, safety properties of smart contracts can be proved in proof assistants like Isabelle/HOL. The \texttt{hevm} project~\cite{hevm} is implemented in Haskell for unit testing and debugging smart contracts, though the EVM implementation is not completed yet. Besides the above formalizations of EVM, there are also implementations of EVM in other languages such as %Javascript~\cite{vmjs}, 
Go~\cite{vmgo} and Ruby~\cite{cryptape}.
%    \item \textbf{Verification of Smart Contracts}

Sergey et al.~\cite{sergey2017concurrent} provided a new perspective between smart contracts and concurrent objects, based on which existing tools and insights for understanding, debugging and verifying concurrent objects can be used on smart contract behaviors. In \cite{luu2016making}, several new security problems were pointed out and a way to enhance the operational semantics of Ethereum was proposed to make smart contracts less vulnerable. Due to the difficulty of correcting the semantics of Ethereum, Luu et al.~\cite{luu2016making} also implemented a symbolic execution tool \texttt{OYENTE} to find security bugs. A framework to analyze and verify both the runtime safety and the functional correctness of Solidity contracts in F* was presented in \cite{bhargavan2016formal}.
%\end{itemize}

\section{Conclusion}\label{Sec: Conclusion}
We propose a framework to enable formal specification, verification and testing towards EVM. In this framework, the formalization of EVM is specified in WhyML, the programming and specification language of Why3. Based on the EVM formalization, automatic SMT solvers and interactive theorem provers provided in Why3 can be employed for verification. The OCaml implementation of EVM is extracted from the WhyML specification and then glued with Rust implementation based on the coupling framework. The coupling framework provides the interaction mechanism between OCaml and Rust, which allows us to perform tests on the new implementation without additional interface implementation.

\subsection*{Acknowledgement}
\noindent This work has been supported by the National Natural Science Foundation of China under grant no. 61772038, 61532019 and 61272160, and the Guangdong Science and Technology Department (Grant no.2018B010107004).
